\Title

### 2022-10-12 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

This raster plot is generated by spiking neurons which are defined on the one hand by the equations governing the evolution of its membrane potential dynamics on its soma and on the other hand by the characterization of the synaptic contacts on its dendritic tree, or on the sensory processes (such as the pixel on the sensor of a camera, or the photoreceptors in the retina). A classical characterization consists in detailing the synaptic weights of each synaptic contact on the dendritic tree, the so-called weight matrix. A more detailed description of these synaptic contacts reveals the added importance of heterogeneous delays, i.e., the precise timing from one afferent neuron's firing to its arrival to the soma, and how this changes the network's dynamics~\citep{izhikevich_polychronization_2006}. In that description, we can parameterize each neuron by the set of tuples defining both the weight and the delay of each synaptic contact. As a consequence, a set of input presynaptic spikes $\event$ will be processed by the dendrites defined by this synaptic set and notably by the respective delays, which will multiplex in time all events.

Formally, let's define a layer of heterogeneous spiking neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$. In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron. A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays. Note that a neuron may contact an afferent neuron with multiple different delays. For each neuron $\postsynaddr$, we thus define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$. This defines the full connectivity of the heterogeneous model. As a consequence, an event stream $\event_\presynaddrspace$ on the presynaptic address space is transformed into a new event stream $\event_\postsynaddrspace$ which is defined by 
$\{ (\presynaddr_\arank = 
 \presynaddr_\ranksyn^\postsynaddr, \timev=\timev_\arank + 
\synapticdelay_\ranksyn^\postsynaddr )_{ \arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]} )
\}$. \note{formaliser le multiplexage}
When post-synaptic neurons are activated on their soma by this discharge motif the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner.

Taking the argument the other way around, one may form a generative model for generic raster plots. Indeed, any spike in the presynaptic address space is generated by sensory neurons (for instance photoreceptors in the retina, sensors in a CMOS chip) or by afferent spiking neurons. In the latter case, these are connected to the spiking cell by a set of weights and delays, whose structure is relatively stable relative to the coding time-scale. When these connections are high and sparsely distributed, this firing will cause a specific temporal motif. Another example is given for the barn owl auditory system: As it hears the sound of a mouse, this sound will generate a specific spiking response in both ears, and specifically, the precise timing between the signal generated by the the left relative to the right ear can be for instance used to determine the position of the prey~\citep{goodman_spike-timing-based_2010}. Overall, these examples show that raster plots may be considered as a mixture of the effects of different elementary causes, and that each event triggers a specific spatio-temporal motif. 
%
\subsubsection{Detecting spiking motifs}
%: Detection model

>>>>

This raster plot is generated by spiking neurons which are defined on the one hand by the equations governing the evolution of its membrane potential dynamics on its soma and on the other hand by the characterization of the synaptic contacts on its dendritic tree, or on the sensory processes (such as the pixel on the sensor of a camera, or the photoreceptors in the retina). A classical characterization consists in detailing the synaptic weights of each synaptic contact on the dendritic tree, the so-called weight matrix. A more detailed description of these synaptic contacts reveals the added importance of heterogeneous delays, i.e., the precise timing from one afferent neuron's firing to its arrival in the soma, and how this changes the network's dynamics~\citep{izhikevich_polychronization_2006}. In that description, we can parameterize each neuron by the set of tuples defining both the weight and the delay of each synaptic contact. As a consequence, a set of input presynaptic spikes $\event$ will be processed by the dendrites defined by this synaptic set and notably by the respective delays, which will multiplex in time all events.

Formally, let's define a layer of heterogeneous spiking neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$. In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron. A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays. Note that a neuron may contact an afferent neuron with multiple different delays. For each neuron $\postsynaddr$, we thus define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$. This defines the full connectivity of the heterogeneous model. As a consequence, an event stream $\event_\presynaddrspace$ on the presynaptic address space is transformed by the synapse $\ranksyn$ of neuron $\postsynaddr$ into a new event stream $\event_\presynaddrspace_\ranksyn$ which is defined by 
$\{ (\presynaddr_\arank = 
 \presynaddr_\ranksyn^\postsynaddr, \timev=\timev_\arank + 
\synapticdelay_\ranksyn^\postsynaddr )_{ \arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]} )
\}$. This new stream of events \note{formaliser le multiplexage}
When post-synaptic neurons are activated on their soma by this spatio-temporal motif the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner.

Taking the argument the other way around, one may form a generative model for generic raster plots. Indeed, any spike in the presynaptic address space is generated by sensory neurons (for instance photoreceptors in the retina, sensors in a CMOS chip) or by afferent spiking neurons. In the latter case, these are connected to the spiking cell by a set of weights and delays, whose structure is stable relatively to the coding timescale. When these connections are high and sparsely distributed, this firing will cause a specific temporal motif. Another example is given for the barn owl auditory system: As it hears the sound of a mouse, this sound will generate a specific spiking response in both ears, and specifically, the precise timing between the signal generated by the left relative to the right ear can be for instance used to determine the position of the prey~\citep{goodman_spike-timing-based_2010}. Overall, these examples show that raster plots may be considered as a mixture of the effects of different elementary causes, and that each event triggers a specific spatio-temporal motif. 
%
\subsubsection{Detecting spiking motifs}
%: Detection model
